---
title: "Data-Driven Education: The Sexuality Blog"
subtitle: "Unveiling the Impact of Sex Education Programs"
author: "Olivia Fountain"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(scales)
library(RColorBrewer)
library(treemapify)
library(plotly)
library(maps)
library(ggwaffle)
library(ggimage)
library(shiny)
```

```{r, include=FALSE, warning=FALSE}
program<-c('Division of Adolescent and School Health', 'Teen Pregnancy Prevention Program', 'Personal Responsibility Education Program', 'Personal Responsibility Education Innovative Strategies', 'Title V State Sexual Risk Avoidance Education', 'General-Departmental Sexual Risk Avoidance Education')
abbreviation<-c('DASH', 'TPPP', 'PREP', 'PREIS', 'SRAE', 'GD SRAE')
funding<-c(34081000, 130000000, 93150000, 10700000, 66390000, 34393408)
ed_type<-c('evidence-based','evidence-based','evidence-based', 'evidence-based',
           'abstinence-only', 'abstinence-only')
program_funding<-data.frame(program, abbreviation, funding, ed_type)
teen_birth_rates<-read.csv("CDC Health Data - Teen Birth Rate 2023-March 2024.csv", header=TRUE)
aids_hiv_sti_rates2021<-read.csv("AIDS, HIV, and STI Rates 2021.csv", header=TRUE)
sex_ed_levels<-read.csv("Sex Education Level Data.csv", header=TRUE)
general_requirements<-read.csv("General Requirements for Sex Ed.csv", header=TRUE)
content_requirements<-read.csv("Content Requirements for Sex Ed.csv", header=TRUE)
sex_ed_funding<-read.csv("Sex Ed Funding by State 2022.csv", header=TRUE)
hiv_sti_data<-read.csv("sti_hiv_data.csv", header=TRUE)
hiv_sti_data$Indicator<-as.factor(hiv_sti_data$Indicator)

general_requirements<-general_requirements%>%
  mutate(sex_ed = ifelse(grepl("X", sex_ed), "required", "not required"))%>%
  mutate(hiv_ed = ifelse(grepl("X", hiv_ed), "required", "not required"))%>%
  mutate(medically_accurate = ifelse(medically_accurate == "" | is.na(medically_accurate), "not required", ifelse(medically_accurate == "X", "required", medically_accurate)))%>%
  mutate(opt_out_allowed = ifelse(opt_out_allowed == "" | is.na(opt_out_allowed), "not required", ifelse(opt_out_allowed == "X", "required", opt_out_allowed)))
categorical_columns<-c("sex_ed", "hiv_ed", "medically_accurate", "opt_out_allowed")
general_requirements<-general_requirements %>%
  mutate_at(vars(categorical_columns), as.factor)

content_requirements<-content_requirements%>%
  mutate(contraception = ifelse(grepl("X", contraception), "required", "not required"))%>%
  mutate(abstinence = case_when(
    abstinence == "Stress" ~ "stressed",
    abstinence == "Cover" ~ "covered",
    is.na(abstinence) | abstinence == "" ~ "not covered",
    TRUE ~ abstinence
  ))%>%
  mutate(sowm = ifelse(grepl("X", sowm), "required", "not required"))%>%
  mutate(sexual_orientation = case_when(
    grepl("Inclusive", sexual_orientation) ~ "inclusive",
    grepl("Negative", sexual_orientation) ~ "exclusive",
    grepl("ᶲ", sexual_orientation) ~ "exclusive",
    grepl("Prohibited", sexual_orientation) ~ "prohibited",
    is.na(sexual_orientation) | sexual_orientation == "" ~ "unknown",
    TRUE ~ sexual_orientation
  ))%>%
  mutate(hiv_condoms = ifelse(grepl("X", hiv_condoms), "required", "not required"))%>%
  mutate(hiv_abstinence = case_when(
    hiv_abstinence == "Stress" ~ "stressed",
    hiv_abstinence == "Cover" ~ "covered",
    is.na(hiv_abstinence) | hiv_abstinence == "" ~ "not covered",
    TRUE ~ hiv_abstinence
  ))
categorical_columns1<-c("contraception", "abstinence", "sowm", "sexual_orientation", "hiv_condoms", "hiv_abstinence")
content_requirements<-content_requirements %>%
  mutate_at(vars(categorical_columns1), as.factor)

program_funding$ed_type<-as.factor(program_funding$ed_type)
program_funding$program<-as.factor(program_funding$program)

teen_birth_rates$percentage_births_in_US <- substr(teen_birth_rates$percentage_births_in_US, 1, nchar(teen_birth_rates$percentage_births_in_US) - 1)
teen_birth_rates$percentage_births_in_US <- as.numeric(teen_birth_rates$percentage_births_in_US)

numeric_columns<-c("aids_classifications", "hiv_diagnoses", "chlamydia", "gonnorhea")
aids_hiv_sti_rates2021<-aids_hiv_sti_rates2021 %>%
  mutate_at(vars(numeric_columns), ~ as.numeric(gsub(",", "", .)))

hiv_sti_data$Cases<-hiv_sti_data$Cases<- as.numeric(gsub(",", "", hiv_sti_data$Cases))

numeric_levels<-c("parents_support_sex_ed", "have_received_sex_ed", "in_elementary", "in_secondary", "in_postsecondary")
sex_ed_levels<-sex_ed_levels %>%
  mutate_at(vars(numeric_levels), ~ as.numeric(gsub("%", "", .)))
sex_ed_levels$region=as.factor(sex_ed_levels$region)
```

```{r, include=FALSE, warning=FALSE}
pregnancies_vs_ed_reqs<-left_join(teen_birth_rates, general_requirements, by="state")
usa_map<-map_data("state")
test_data<-pregnancies_vs_ed_reqs
test_data$births<-as.numeric(test_data$births)
test_data$state<-tolower(test_data$state)
data_map<-usa_map %>%
  left_join(test_data, by=c("region" = "state"))
data<-pregnancies_vs_ed_reqs
funding_and_diagnoses<-left_join(hiv_sti_data, sex_ed_funding, by="state")
```

In the United States, the topic of sexual education in schools has long been divisive, with supporters of both evidence-based and abstinence-only instruction. These two approaches to sexual education tend to reflect opposing ideologies about educating the youth about relationships, contraception, and sexual health. Abstinence-only education emphasizes waiting until marriage to engage in sexual activity, or not engaging whatsoever as the only way to avoid getting pregnant and contracting STIs. Advocates contend that teaching youth only about abstinence is in line with moral and religious principles and encourages beneficial behavioral results. Such instruction often falls short in its ability to delay sexual initiation or minimize risky sexual behaviors. In contrast, evidence-based and comprehensive sexual education places a strong emphasis on accurate, age-appropriate knowledge on a variety of sexual health-related subjects, such as consent, STI prevention, healthy relationships, contraception, and sexual orientation. Supporters argue that evidence-based sexual education gives youth the information and abilities they need to make wise choices regarding their sexual health and well being. It is essential to discern the differences between evidence-based sexual education and abstinence-only in order to influence educational practices, inform public policy, and improve the health of adolescents. 

Federal funding for sex education programs in the 2022 fiscal year has been controversial, with more than 25% of total funding going toward programs that promote abstinence-only education. The values in the following visualization are inclusive of all funding nation-wide. Studies from prestigious research programs such as Harvard’s Human Rights Institution have concluded through endless surveys and studies that abstinence-only education is both ineffective and could potentially be harmful to young people (Kay & Jackson, 2008). Although the majority of federal funding was granted to evidence-based programs, continuing to invest over 100 million dollars into abstinence-only programs given the evidence of its inefficacy only increases the urge to provoke lawmakers into considering defunding abstinence-only programs and increasing the amount of support for evidence-based and comprehensive programs.
```{r, echo=FALSE}
pie_data<-program_funding
funding_plot <- plot_ly(pie_data)%>%
  add_pie(
  data = pie_data,
  values = ~funding,
  labels=~abbreviation,
  textposition="outside",
  textinfo="label+percent",
  hoverlabel="text",
  text=~paste(program, '<br>', ed_type),
  type = "pie",
  marker=list(colors=c('#e3d2e3',"thistle", "#dfcbdf",'#dbc5db', "#801818","#8c2f2f")),
  pull = 0.01,
  hole=0.3)%>%
  layout(title = list(text="Sex Education Federal Program Funding", y=0.9),
        images = list(
      source = base64enc::dataURI(file = "legend.png"),
      x = 0.8, y = 0.8, 
      sizex = 0.3, sizey = 0.2,
      xref = "paper", yref = "paper", 
      xanchor = "left", yanchor = "bottom"
    ), showlegend=FALSE,  margin = list(t = 75, b=75, l=80, r=80)
)
funding_plot
```


Iowa, Minnesota, New York, and Pennsylvania all receive funding for abstinence-only education even though their sex education curricula does not require abstinence-only coverage. This discrepancy highlights an imbalance between the distribution of funds and state-level sex education policy mandates. Even in the absence of legal requirements, states may choose to subsidize programs that educate people about abstinence as a way to address perceived moral or cultural norms. In order to make sure that funding decisions take into consideration the requirements and realities of adolescent sex education, there is a need for increased transparency and accountability as the conversation around this funding continues.

```{r, echo=FALSE, warning=FALSE}
abstinence_education_vs_funding<-left_join(sex_ed_funding, content_requirements, by="state")
abstinence_education_vs_funding<-abstinence_education_vs_funding%>%
  select(state, evidence_based_only, abstinence_only, abstinence)%>%
  filter(abstinence == 'not covered')
abstinence_education_vs_funding$absdollar<-paste0("$", format(abstinence_education_vs_funding$abstinence_only, big.mark = ",", scientific = FALSE))
addUnits <- function(n) {
  labels <- ifelse(n < 1000, n,
                   ifelse(n < 1e6, paste0(round(n/1e3), 'K'), 
                          ifelse(n < 1e9, paste0(round(n/1e6), 'M'), 'n')))
  return(labels)
}
ggplot(abstinence_education_vs_funding, aes(x = state, y=abstinence_only)) +
  geom_bar(stat = "identity", fill="#801818") +
  labs(title = "States Receiving Abstinence-Only Program Funding That Do Not Require Abstinence Coverage",
       x = "State",
       y = "Funding") +
  theme_minimal() + scale_y_continuous(labels = addUnits,  breaks = pretty_breaks(n = 7)) +
  theme(plot.title=element_text(size=10, hjust = 0.5)) +
  geom_text(aes(label=absdollar, vjust = -0.8), size = 3, color = "black")
```


The most recent statistics on adolescent pregnancy rates from January 2023 to March 2024 can provide useful insight into how sex education laws may affect teen pregnancy. Given the following visualization tool, adolescent pregnancy rates are generally lower across the midwest, and tend to increase in the south and western regions of the US.  It is interesting that although four out of the top five states with the highest count of adolescent pregnancies require sex education, over half of the United States does not. This application allows for a more detailed overview of both the adolescent pregnancy rates and sex education requirements by state. When reviewing this data, one of the most discouraging findings can be found in the ‘Sex-Ed Requirements’ map, where over 30% of the United States does not even require sex education to be medically accurate. As a further clarification, the ‘Sex Education’ requirement type includes both evidence-based and abstinence-only programs. 

```{r, echo=FALSE, out.extra="data-external=1"}
knitr::include_app(url = "https://ogf9uhy.shinyapps.io/AdolescentPregnancyRate/", height="800px")
```

A much more serious discussion revolves around the impact of AIDs and STIs as a threat to public health, with state-by-state variations in infection rates.  This animated visualization compares state-by-state infection rates of HIV, AIDs classifications, Gonorrhea, and Chlamydia from 2021 by the amount of funding each state received for both evidence-based and abstinence-only education programs. In order to better scale the data, all axes have been log transformed, but the hover text provides more details into the exact amount of funding and diagnoses per state. Given these two visuals, it is apparent that states receiving larger amounts of funding for abstinence-only programs still have considerably high STI rates. It is important to note that these findings can also be influenced by population size and other demographic factors, but a core insight into these plots is that the trend of STI rates is still increasing, despite having  more funding and support for abstinence-only programs. 
```{r, warning=FALSE, echo=FALSE}
funding_and_diagnoses$ebdollar<-paste0("$", format(funding_and_diagnoses$evidence_based_only, big.mark = ",", scientific = FALSE))
funding_and_diagnoses$absdollar<-paste0("$", format(funding_and_diagnoses$abstinence_only, big.mark = ",", scientific = FALSE))
funding_and_diagnoses$Indicator<-reorder(funding_and_diagnoses$Indicator, desc(funding_and_diagnoses$Indicator))

evidence_based<-funding_and_diagnoses %>%
  plot_ly(x = ~log10(evidence_based_only), y = ~log10(Cases), text=~paste(state,'<br>', "Funding: ", ebdollar, '<br>', "2021 Diagnoses : ", Cases), hoverinfo="text", mode="markers", marker=list(size=11, color="thistle", opacity=0.8))%>%
  add_markers(data = funding_and_diagnoses, frame=~Indicator) %>%
  hide_legend() %>%
  animation_opts(frame = 1000, transition = 0, redraw = FALSE)%>%
    layout(xaxis=list(title="Funding"), yaxis=list(title="Cases"))

abstinence_only<-funding_and_diagnoses %>%
  plot_ly(x = ~log10(abstinence_only), y = ~log10(Cases), text=~paste(state,'<br>', "Funding: ", absdollar, '<br>', "2021 Diagnoses : ", Cases), hoverinfo="text", mode="markers", marker=list(size=11, color="royalblue", opacity=0.6))%>%
  add_markers(data = funding_and_diagnoses, frame=~Indicator) %>%
  hide_legend() %>%
  animation_opts(frame = 1000, transition = 0, redraw = FALSE)%>%
  layout(xaxis=list(title="Funding"), yaxis=list(title="Cases"))
annotations = list( 
  list( 
    x = 0.2,  
    y = 1.0,  
    text = "Evidence Based Programs",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),  
  list( 
    x = 0.8,  
    y = 1,  
    text = "Abstinence-Only Programs",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ))
both_plots<- subplot(evidence_based, abstinence_only)%>%layout(annotations = annotations, yaxis=list(title="Funding"), xaxis=list(title="Diagnoses"), title = "")
both_plots
```


In the battle against HIV transmissions, lowering infection rates through education on condom usage and HIV preventive techniques is essential. A thorough education on HIV/AIDS equips people with information on risk factors, preventative strategies, and mechanisms of transmission. The following visualization represents the top 15 states with the highest number of HIV diagnoses from the CDC in 2021. Interestingly, over half of these states do not require any instruction  on condom usage in the context of HIV prevention education. States that stress abstinence-only sex education often provide those same methods when addressing HIV prevention, which is the opposite for evidence-based programs. Pregnancy and STI prevention methods (other than abstinence) such as condom usage and contraception are widely taught across all different forms of evidence-based programs.  These programs help to promote a culture of prevention and accountability, which can eventually improve public health outcomes.
```{r, include=FALSE, warning=FALSE}
hiv_requirements<-left_join(general_requirements, content_requirements, by="state")
hiv_requirements<-left_join(hiv_requirements, aids_hiv_sti_rates2021, by="state")
hiv_required<-hiv_requirements%>%
  select(state, hiv_ed, hiv_condoms, aids_classifications, hiv_diagnoses)%>%
  filter(hiv_ed== 'required' & hiv_condoms == 'required')
hiv_not_required<-hiv_requirements%>%
  select(state, hiv_ed, hiv_condoms, aids_classifications, hiv_diagnoses)%>%
  filter(hiv_ed== 'not required' & hiv_condoms == 'not required')
```

```{r, echo=FALSE, warning=FALSE, fig.width=7, fig.height=5}
hiv_requirements<-hiv_requirements[order(-hiv_requirements$hiv_diagnoses),]
top_15<-head(hiv_requirements, 15)
tree_map<-ggplot(data=top_15, aes(area=hiv_diagnoses, fill=hiv_condoms, label = state)) +
  geom_treemap(alpha=0.8) +
  geom_treemap_text(place = "center", color = "black", size = 12) +
  labs(title = "Top 15 States Ranking Highest in HIV Diagnoses", fill ="Condom Use Education for HIV Prevention") +
  theme(plot.title = element_text(hjust = 0.5, size=10), legend.title=element_text(size=10)) +
  scale_fill_manual(values = c("royalblue3", "thistle3")) 
tree_map
```


In the US, there is a great deal of variation in the context of sex education. While some states require complete, evidence-based programs, others place a higher priority on teaching focused solely on abstinence. Even states that do not require any abstinence-only coverage still stress some of the fundamental principles of that instruction, such as sex only occurring when married. To elaborate on this perspective, the following visualization tool shows the count of states that provide coverage or require certain curricula in sex education programs. Referring back to the earlier visualization tool for sex education requirements,  only 22 states require sex education, yet over 35 states either cover or stress abstinence-only instruction in their curriculum.  Abstinence-only programs usually omit information on contraception and safer sex practices, which can deprive adolescents of the knowledge and skills they need to make educated decisions about their sexual health.

```{r, echo=FALSE, out.extra="data-external=2"}
knitr::include_app(url = "https://ogf9uhy.shinyapps.io/SexEdContentRequirements", height="800px")
```

To further the conversation about sex education exposure, the following visualization represents the findings from a series of qualitative interview surveys conducted by the Bedbible Research Center that has ranged from 1995 to 2023, and is still being continuously updated (Bedbible Research Center, 2022). The survey includes the percentage of respondent answers about their sex education exposure, aggregated by state. Since observing some differences between adolescent pregnancy rates and program requirements across groups of states, it may be of interest to compare these responses by region. Given this visual, it appears that midwestern states had the highest percentage of overall sex education exposure, while the southern and western regions had the lowest. This distinction could be of use in supporting more targeted interventions and tailored educational approaches to address gaps in sex education access and promote evidence-based education across all regions of the United States.
```{r, include=FALSE, warning=FALSE}
midwest<-sex_ed_levels%>%
  filter(region=='Midwest')%>%
  summarise(parents = median(parents_support_sex_ed),
            have_rcvd = median(have_received_sex_ed),
            elementary = median(in_elementary),
            secondary = median(in_secondary),
            postsec = median(in_postsecondary))
south<-sex_ed_levels%>%
  filter(region=='South')%>%
  summarise(parents = median(parents_support_sex_ed),
            have_rcvd = median(have_received_sex_ed),
            elementary = median(in_elementary),
            secondary = median(in_secondary),
            postsec = median(in_postsecondary))
west<-sex_ed_levels%>%
  filter(region=='West')%>%
  summarise(parents = median(parents_support_sex_ed),
            have_rcvd = median(have_received_sex_ed),
            elementary = median(in_elementary),
            secondary = median(in_secondary),
            postsec = median(in_postsecondary))
northeast<-sex_ed_levels%>%
  filter(region=='Northeast')%>%
  summarise(parents = median(parents_support_sex_ed),
            have_rcvd = median(have_received_sex_ed),
            elementary = median(in_elementary),
            secondary = median(in_secondary),
            postsec = median(in_postsecondary))
levels_by_region<-bind_rows(midwest, south, west, northeast)
levels_by_region<-levels_by_region%>%
  mutate(regions=c("Midwest", "South", "West", "Northeast"))
levels_by_region<-read.csv("median_responses_region.csv")
levels_by_region$level<-as.factor(levels_by_region$level)
levels_by_region$region<-as.factor(levels_by_region$region)
```

```{r, echo=FALSE, warning=FALSE}
sex_ed_exposure<-ggplot(levels_by_region, aes(x = region, y=median_percentage, fill=level)) +
  geom_bar(position=position_dodge(width=0.7), stat="identity", width=0.5) +
  labs(title = "Sex Education Exposure Across Regions", x = "Region", y = "Median Percentage of Responses", fill = "Received Sex Education") +
  theme_minimal() +
  scale_y_continuous(labels = function(x) paste0(x, "%"), breaks = pretty_breaks(n = 7)) +
  scale_fill_manual(values=c("tan2", "royalblue3", "thistle", "#801818")) +
    theme(plot.title = element_text(hjust = 0.5, size=10))
sex_ed_exposure
```


Lastly, it is crucial to reflect on the nationwide percentage of LGBTQ+ youth in the United States and how much disparity is often reflected in sexual orientation education policies. 
According to the CDC, the 2020 Youth Risk Behavior Surveillance Survey estimated that 9.5% of the youth population between the ages of 13-17 identified within the LGBTQ+ community (Williams Institute, 2020). Currently, only 11 states have endorsed the inclusion of sexual orientation education in their programs, leaving many LGBTQ+ youth without access to comprehensive information about sexual orientation, identity, and related health issues. This lack of coverage in the majority of states, regardless of sex education instruction, underscores the importance of addressing disparities in sexual orientation education to support the well-being of LGBTQ+ youth nationwide. Additionally, federal guidance has prohibited abstinence-only programs from teaching about sexual orientation or gender identity (Ott et al, 2007). By advocating for inclusive sexual education policies and ensuring that all young people have access to accurate and affirming information, policymakers can promote a safer and more supportive environment for LGBTQ+ youth to thrive.
```{r, echo=FALSE, warning=FALSE, fig.align='left'}
## https://stackoverflow.com/questions/69133845/visualize-quantities-by-icons-using-r
inclusive_education<-table(content_requirements$sexual_orientation)
icon_plot_data<-tibble(state= rep("inclusive", 11) %>% c(rep("exclusive", 39)))
waffle_data <- waffle_iron(icon_plot_data, aes_d(group = state), rows = 5) %>% mutate(label = ifelse(icon_plot_data$state == "inclusive", "rainbow circle.png", "grey circle.png"))

icon_array<-ggplot(waffle_data, aes(x, y)) + 
  geom_image(aes(image = label), size = 0.12) +  # Adjust size as needed
  coord_equal() + 
  theme_void() +  # Remove background and axis
  expand_limits(x = c(0, 4), y = c(0, 6)) +  # Adjust plot limits
  labs(x = "", y = "") +
    theme(plot.title = element_text(size = 10, hjust=0.5))
icon_array
```
In conclusion, the analysis between the impacts of evidence-based and abstinence-only sex education highlights how vital it is to offer complete and inclusive sexual education. Adolescent sexual health can be routinely improved by comprehensive sex education, which covers topics including consent, STI prevention, contraception, and sexual orientation. As these conversations progress, it is critical that decision-makers give evidence-based strategies top priority and appropriate funding to guarantee that all young people have access to the information and assistance they require to make responsible decisions regarding their sexual health and wellbeing.

To guarantee that everyone, regardless of skills or disabilities, can access and benefit from these resources, accessibility issues must be addressed in addition to the content of sex education programs. Different kinds of color deficiencies can be accommodated in this blog post by the visualizations that have a discrete color scale with one overall hue ranging from light to dark. Furthermore, the addition of blue and orange to plots with more than two colors can make them more accessible to those who have red-green color vision impairments, encouraging inclusivity and making sure that everyone can interact with and comprehend the content being displayed.

### Resources
Bedbible Research Center. (2022, December 6). Sex Education Statistics - The state of sexual education (+Dataset). Bedbible.com. https://bedbible.com/sex-education-statistics/

Centers for Disease Control and Prevention. (n.d.). AtlasPlus - HIV, Viral Hepatitis, STD, TB, and Social Determinants of Health. Gis.cdc.gov. https://gis.cdc.gov/grasp/nchhstpatlas/tables.html

Guttmacher Institute. (2023, September 1). Sex and HIV education. Guttmacher Institute. https://www.guttmacher.org/state-policy/explore/sex-and-hiv-education

Kaiser Family Foundation. (2018, June 1). Abstinence Education Programs: Definition, Funding, and Impact on Teen Sexual Behavior. The Henry J. Kaiser Family Foundation. https://www.kff.org/womens-health-policy/fact-sheet/abstinence-education-programs-definition-funding-and-impact-on-teen-sexual-behavior/

Kay, J., & Jackson, A. (2008). How Abstinence-Only Programs Harm Women and Girls Sex, Lies & Stereotypes. https://hrp.law.harvard.edu/wp-content/uploads/2013/03/sexlies_stereotypes2008.pdf

Ott, M. A., & Santelli, J. S. (2007). Abstinence and abstinence-only education. Current opinion in obstetrics & gynecology, 19(5), 446–452. https://doi.org/10.1097/GCO.0b013e3282efdc0b

Sex Ed for Social Change (SIECUS). (2022). Federal Funding Overview: Fiscal Year 2022. https://siecus.org/wp-content/uploads/2022/05/FY22-Federal-Funding-Overview.pdf

Williams Institute. (2020, September). LGBT Youth Population in the United States. Williams Institute. https://williamsinstitute.law.ucla.edu/publications/lgbt-youth-pop-us/

World Population Review. (2021). Abstinence Only Education States 2021. Worldpopulationreview.com. https://worldpopulationreview.com/state-rankings/abstinence-only-education-states